Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2020-08-07 Manuscript/HI_Model_ms.Rmd
updated manuscript files

### *States and transitions*
We developed a high and a low land-use scenario, with transition pathways modified from @daniel_state-and-transition_2016. Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to developed. Although several state types had urbanization transition pathways, but there was no transition pathway from developed to any other state type. Water and wetland state types remained static throughout the simulation period.

 Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaii [@kim_population_2018]. For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land-use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016]. In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion of *Eucalyptus* spp. plantations to agriculture. The remaining 40% were rotation harvests with replanting of *Eucalyptus* spp.

The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (Figure S2) to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1). The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017. Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019, while the high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (Fig. S2a).

### *Carbon stocks and flows*

>>>>

### *States and transitions*
We developed two land-use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016. Transitions between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, tree plantation, and shrubland state types each had multiple transition pathways, while the barren state type could only transition to developed (i.e., urbanization). There was no transition pathway out of an urbanized (developed) state. Water and wetland state types remained static throughout the simulation period.

 Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaii [@kim_population_2018]. For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land-use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016]. In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.

The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (Figure S2). We used this new Hawai`i fire database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1) for the years 1999-2019. The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017. Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019, while the high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (Fig. S2a).

### *Carbon stocks and flows*
 The fate of continuous carbon stocks was tracked for each simulation cell based on a suite of continuous flows (i.e., carbon fluxes) specifying the rates of change in these carbon stocks over time [@daniel_integrating_2018; @sleeter_effects_2019].
